Does It Scale? Static Output Approximation of PHP Web Applications Using Symbolic Execution

### 2017-01-14 paper.tex
4.0 4.1 and 4.2 refactored

\definecolor{lightgrey}{HTML}{FFFFFC}

@@

\subsection{Experience Report}
We identified trade-offs between accuracy and performance that may
cause the approximation to be impractical, yet possible. Additionally, we
detected limitations impeding symbolic execution conceptually: For certain
language features, concrete information rather than symbolic values may be required to
actually be executed properly as this defective execution can impact subsequent
parts of the program which may become inaccessible and may be missed.

@@

\section{Evaluation}
As already stated in the previous section, we encountered sevaral problems
during the development of the symbolic interpreter. Therefore this section
describes our methods to evaluate practicality of the output approximation, and
possible explanations.

\subsection{Measuring Approximation Success} \label{heuristic}
For the evaluation of our output approximation we require a ground truth output
model to compare our approximation against. Since output literals are embedded
in server-side code and scattered, it is nontrivial to determine whether a
string literal is part of the ouput or not. So, in the absence of ground truth
for the approximated output we are bound to choose a \emph{heuristic approach}
in order to measure the accuracy of our aproximation. We define a string
literal to be an \emph{output candidate} if there exists an execution path
reaching the string literal and evantually passing it to an output-generating
statement (e.g., \texttt{echo} or \texttt{print}).

We approximate output candidates as those string literals containing the
characters \texttt{<}, \texttt{>}, or both since we expect output to contain
HTML tags. We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from the entire corpus.
We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 96 percent of the string literals are classified correctly. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output, actually were responsive to the classifier. We attempted
to increase recall by looking for further distinctive properties to build a
classifier from, but could not do so without decreasing precision. In spite of
missing half of the output candidates, we decided to use a simple, yet
distinctive classifier.

\subsection{Experiment Setup} \label{sec:experiment_setup}
For the evaluation we symbolically executed a selection of PHP systems. We
selected a corpus of twelve PHP systems with regard to system size and recency
as the corpus of the case study for \cite{Nguyen:2014:BCG:2635868.2635928} only
contained small-scale systems that are not maintained any more. The full list
of PHP systems is shown in Table \ref{corpus}. Our selection of PHP systems
includes%
\begin{itemize}
	\item four small-scale systems that selected from the previously mentioned case
	study in order to compare our results with the previous symbolic execution
	engine \cite{Nguyen:2014:BCG:2635868.2635928},

	\item three recent large-scale systems selected from the case study corpus of
	\cite{Hills:2013:ESP:2483760.2483786}, and

	\item four small-scale systems selected from a list of recent Content
	Management Systems \cite{codegeekz}.
\end{itemize}

As for for the last two cases, we limited our selection since the parser we
used did not support all language features used. For the experimental evaluation we symbolically executed per system each script file or entry point (files with a \texttt{.php}, \texttt{.inc}, \texttt{.bit} or \texttt{.module} extension). All following measurements are cumulated per system for each entry point.

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{4cm}rrrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}
	\\
	\midrule
	AddressBook & 8.2.5.2 &  & 51,907 & 239  & 1009 & 186\\
	SchoolMate & 1.5.4 &  & 8,118 & 65  & 853 & 88\\
	TimeClock & 1.04 &  & 20,800 & 63  & 7920 & 306\\
	WebChess & 1.0.0 &  & 5,219 & 28  & 470 & 56\\
	\midrule
	Drupal & 7.5.0 & CMS & 52,464 & 125  & 3569 & 749\\
	phpBB & 3.1.9 &  & 327,371 & 1,398  & 3606 & 206\\
	phpMyAdmin & 4.6.3 &  & 303,582 & 871  & 7103 & 571\\
	\midrule
	Anchor & 0.12.1 &  & 15,054 & 201 & 987 & 32\\
	Kirby & XXX & CMS & XXX & XXX  & 654 & 23\\
	Automad & XXX & CMS & XXX & XXX  & 655 & 8\\
	Monstra & XXX & CMS & XXX & XXX  & 1934 & 48\\
	Nibbleblog & XXX & CMS & XXX & XXX  & 1013 & 28\\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

@@

\subsection{Understanding Limitations}
As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
coverage metrics were poor for large-scale and small-scale/recent systems. To
understand what output candidates we missed and why, we further investigated
our approximation results and conducted three more measurements.

@@

\section{Related Work}
\cite{Hills:2013:ESP:2483760.2483786}
\cite{wassermann2007sound,minamide2005static}
\textcolor{blue}{\blindtext\blindtext}

>>>>

\definecolor{lightgrey}{HTML}{CCFFEE}

@@

\subsection{Experience Report}\label{sec:experience_report}
We identified trade-offs between accuracy and performance that may
cause the approximation to be impractical, yet possible. Additionally, we
detected limitations impeding symbolic execution conceptually: For certain
language features, concrete information rather than symbolic values may be required to
actually be executed properly as this defective execution can impact subsequent
parts of the program which may become inaccessible and may be missed.

@@

\section{Evaluation}
As already stated in the previous Section \ref{sec:experience_report}, we
encountered several problems during the implementation of our symbolic
execution engine for \sf{WordPress}. The main items of critique of the symbolic
execution engine and tools described in previous work
\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
are the missing support for widely-used language features, and the evaluation
only using small-scale systems. Although evaluating a tool for a small systems can be a valuable proof-of-concept, it does not necessarily cover the question of whether a tool is practical on a larger-scale. Therefore in the following we present our
methods to evaluate practicality of our symbolic output approximation.
Moreover, we investigate to which extent the conceptual limitations encountered
with \sf{WordPress} apply for a wider choice of PHP systems.

\subsection{Experiment Setup} \label{sec:experiment_setup}
We have encountered several issues for \sf{WordPress} that conceptually limit
the scalability of symbolic execution. Since our observations so far only considers
one example system an, we evaluate our symbolic execution engine for a wider
range of PHP systems with regard to the following questions. The metrics to
approach those questions are explained in Sections
\ref{HowAccurateIsOurApproximation} and \ref{sec:understanding_limitations}.

First, the evaluation of the output approximation using the symbolic execution
engine, \sf{Symex}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928} has
shown high code coverage for small systems. As our symbolic execution
re-implements \sf{Symex} and extends with support for additional language
features, we ask whether we can replicate high coverage results at least on a small-scale. Therefore, our PHP
corpus contains four small-scale systems used in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.

Second, we ask whether system size and usage of modern language features are
factors affecting the practicality of our symbolic execution engine. We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
footnote}, did not support a number of language features. The full list of PHP
systems for that we evaluate our symbolic execution engine along with
descriptive code metrics can be found in Table \ref{corpus}.

\begin{table*}[t]
\centering 
	\begin{tabular}{lrp{4cm}rrrr}
	\toprule
	\textbf{System} & \textbf{Version} & \textbf{Classification} & \textbf{SLOC} &
	\textbf{\#files} & \textbf{\#OCs} & \textbf{\#includes}\\
	\midrule
	AddressBook & 8.2.5.2 &  & 51,907 & 239  & 1009 & 186\\
	SchoolMate & 1.5.4 &  & 8,118 & 65  & 853 & 88\\
	TimeClock & 1.04 &  & 20,800 & 63  & 7920 & 306\\
	WebChess & 1.0.0 &  & 5,219 & 28  & 470 & 56\\
	\midrule
	Drupal & 7.5.0 & CMS & 52,464 & 125  & 3569 & 749\\
	phpBB & 3.1.9 &  & 327,371 & 1,398  & 3606 & 206\\
	phpMyAdmin & 4.6.3 &  & 303,582 & 871  & 7103 & 571\\
	\midrule
	Anchor & 0.12.1 &  & 15,054 & 201 & 987 & 32\\
	Kirby & XXX & CMS & XXX & XXX  & 654 & 23\\
	Automad & XXX & CMS & XXX & XXX  & 655 & 8\\
	Monstra & XXX & CMS & XXX & XXX  & 1934 & 48\\
	Nibbleblog & XXX & CMS & XXX & XXX  & 1013 & 28\\
	\bottomrule
	\end{tabular}
	\caption{Corpus of twelve PHP systems. The file count includes files with a .php,
	.inc, .bit or .module extension.}
	\label{corpus}
\end{table*}

\subsection{Measuring Approximation Success} \label{heuristic}
In order to accurately answer the question how good our symbolic execution
engine approximates client page output, we would require knowledge of all
possible variants to compare our approximation against. Since we do not have
any ground truth information alike, we approach this demand by using a
heuristic. Rather than having knowledge about all client page output variants
we say our approximation is sound if all string literals (which are embedded in
and  scattered across server-side code) that may eventually become part of any
client page output variant (in the following referred to as \emph{output
candidate}) is contained in our symbolic output approximation.
It is non-trivial to determine whether a string literal is an output candidate,
so we heuristically classify string literals as output candidates if they
contain the characters \tt{<}, \tt{>}, or both since we expect output to contain
HTML tags.

We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from our corpus of PHP systems (see Table
\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 96 percent of the string literals are classified correctly. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output candidates, actually were responsive to the classifier. We
attempted to increase recall by looking for further distinctive properties to build a
classifier from, but could not do so without decreasing precision. 
\textcolor{blue}{In spite of missing half of the output candidates, we decided to use a simple, yet
precise and distinctive classifier.}

@@

\subsection{Understanding Limitations} \label{sec:understanding_limitations}
As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
coverage metrics were poor for large-scale and small-scale/recent systems. To
understand what output candidates we missed and why, we further investigated
our approximation results and conducted three more measurements.

@@

\section{Related Work} \label{sec:related_work}
\cite{Hills:2013:ESP:2483760.2483786}
\cite{wassermann2007sound,minamide2005static}
\textcolor{blue}{\blindtext\blindtext}
